---
title: 'Replication Code for ''Is Random Forest Really Better than Logistic Regression
  for Predicting Civil War Onsets?'' '
output:
  pdf_document: default
  html_notebook: default
---



```{r Install and Load the packages needed for the analysis, results = 'hide', message = FALSE}
source("packages.R")
```

The packages needed for the analysis will be automatically installed (if they aren't installed yet) and loaded.

```{r Detect number of cores and register number of cores for usage in parallel computation, results = 'hide'}
registerDoMC(cores = detectCores()-1) 
```

```{r Clean Workspace and read Muchlinski et al replication data}
rm(list=ls())

data <- read.csv(file="SambnisImp.csv") 
```

# Data Pre-processing

```{r Select same variables as in Muchlinski et al. and convert DV to factor}
df <- data[,c("warstds", "ager", "agexp", "anoc", "army85", "autch98", "auto4",
              "autonomy", "avgnabo", "centpol3", "coldwar", "decade1", "decade2",
              "decade3", "decade4", "dem", "dem4", "demch98", "dlang", "drel",
              "durable", "ef", "ef2", "ehet", "elfo", "elfo2", "etdo4590",
              "expgdp", "exrec", "fedpol3", "fuelexp", "gdpgrowth", "geo1", "geo2",
              "geo34", "geo57", "geo69", "geo8", "illiteracy", "incumb", "infant",
              "inst", "inst3", "life", "lmtnest", "ln_gdpen", "lpopns", "major", 
              "manuexp", "milper", "mirps0", "mirps1", "mirps2", "mirps3", "nat_war", "ncontig",
              "nmgdp", "nmdp4_alt", "numlang", "nwstate", "oil", "p4mchg",
              "parcomp", "parreg", "part", "partfree", "plural", "plurrel",
              "pol4", "pol4m", "pol4sq", "polch98", "polcomp", "popdense",
              "presi", "pri", "proxregc", "ptime", "reg", "regd4_alt", "relfrac", 
              "seceduc","second", "semipol3", "sip2", "sxpnew", "sxpsq", 
              "tnatwar", "trade", "warhist", "xconst")]

# Converting DV into Factor with names for Caret Library

df$warstds<-factor(
  df$warstds,
  levels=c(0,1),
  labels=c("peace", "war"))
```

```{r Split dataset into training and test data}
dfTrain <- df[data$year < 1990,]
dfTest <- df[data$year >= 1990,]
```

# Training the models

All models are trained using the `caret` R-package.

## Setting up train control

```{r Set training control for caret models}
# create a list of seesd, here change the seed for each resampling
set.seed(221)

# length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#3 is the number of tuning parameters, mtry for rf
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 3)

#for the final model
set.seed(221)
seeds[[11]]<-sample.int(1000, 1)

# Create folds
folds <- 10
cvIndex <- createFolds(factor(dfTrain$warstds), folds, returnTrain = T)

tc <- trainControl(method = "cv", 
                   number = folds,
                   summaryFunction = twoClassSummary, 
                   classProb = TRUE,
                   savePredictions = TRUE,
                   seeds = seeds,
                   index = cvIndex
)
```

## The models with the Fearon/Laitin variables.

First, we train the Logit and Random Forest model using the eleven Fearon/Laitin variables. And have a brief first look at the results.

### Logit
```{r The Fearon/Laitin Logit model, results = 'hide'}
modelFLlogit <- train(warstds ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil 
                      + nwstate + inst3 + pol4 + ef + relfrac, 
                      metric = "ROC", method = "glm", family = "binomial",
                      trControl = tc, data = dfTrain
                      )

summary(modelFLlogit) # Not in paper

modelFLlogit # Not in paper

confusionMatrix(modelFLlogit, norm = "average") # Not in paper
```

### Random Forest
```{r The Random Forest model using the Fearon/Laitin variables, results = 'hide'}
modelFLrf <- train(warstds ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil 
                   + nwstate + inst3 + pol4 + ef + relfrac, 
                   metric = "ROC", method = "rf",
                   trControl = tc, data = dfTrain,
                   keep.inbag = TRUE
                   )

modelFLrf # Not in paper

confusionMatrix(modelFLrf, norm = "average") # Not in paper
```

## The models using all 90 variables in the dataset.

### Logit
```{r The all variables Logit model, results = 'hide'}
modelAlllogit <- train(warstds ~ ., 
                      metric = "ROC", method = "glm", family = "binomial",
                      trControl = tc, data = dfTrain
)

summary(modelAlllogit) # Not in paper

modelAlllogit # Not in paper

confusionMatrix(modelAlllogit, norm = "average") # Not in paper
```

### Random Forest
```{r, results = 'hide'}
modelAllrf <- train(warstds ~ ., 
                   metric = "ROC", method = "rf",
                   trControl = tc, data = dfTrain,
                   keep.inbag = TRUE
)

modelAllrf # Not in paper

confusionMatrix(modelAllrf, norm = "average") # Not in paper
```

## The plots and confusion matrices to compare the models

### Entries for Table 1: Confusion Matrices for the models at a threshold for positive prediction of 0.5
```{r Entries for Table 1}
table(pred = predict(modelFLlogit, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelFLrf, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAlllogit, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAllrf, dfTest, type = "raw"), obs = dfTest$warstds)
```

### Figure 1: Separation Plots for the four models
```{r Separation Plots for the models in the paper, fig.align = 'center', fig.height = 4, fig.width = 7}

separationplot(predict(modelFLlogit, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "Fearon and Laitin Variables (2003) Logit", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelFLrf, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "Fearon and Laitin Variables (2003) Random Forest", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelAlllogit, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "All Variables Logit", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelAllrf, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "All Variables Random Forest", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)
```

### Figure 2: ROC curves for the four models
```{r ROC curves for the models in the paper}
plot.roc(dfTest$warstds,
         predict(modelFLlogit, dfTest, type = "prob")$war,
         col = "black", las = 1, xlim = c(1,0), bty = "n", 
         main = "Out-of-sample ROC curves")

plot.roc(dfTest$warstds,
         predict(modelFLrf, dfTest, type = "prob")$war,
         add = T,
         col = "black", lty = "dashed")


plot.roc(dfTest$warstds,
         predict(modelAlllogit, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "solid")

plot.roc(dfTest$warstds,
         predict(modelAllrf, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "dashed")
legend(0.7, 0.25, 
       c(paste("Fearon and Laitin (2003) Logit, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelFLlogit, dfTest, type = "prob")$war)$auc),2)), 
         paste("Fearon and Laitin (2003) Random Forest, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelFLrf, dfTest, type = "prob")$war)$auc),2)), 
         paste("All Variables Logit, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelAlllogit, dfTest, type = "prob")$war)$auc),2)),  
         paste("All Variables Random Forest, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelAllrf, dfTest, type = "prob")$war)$auc),2))), 
       lty=c("solid", "dashed", "solid", "dashed"), 
       col = c("black","black", "grey 50", "grey 50"), bty="n", 
       cex = .75)


```
```{r Test for difference in ROCs and CIs for AUCs, results = 'hide'}

rocFLlogit <- roc(dfTest$warstds,predict(modelFLlogit, dfTest, type = "prob")$war)
rocFLrf <- roc(dfTest$warstds,predict(modelFLrf, dfTest, type = "prob")$war)

rocAlllogit <- roc(dfTest$warstds,predict(modelAlllogit, dfTest, type = "prob")$war)
rocAllrf <- roc(dfTest$warstds,predict(modelAllrf, dfTest, type = "prob")$war)

set.seed(221)
roc.test(rocAlllogit, rocAllrf, method = "bootstrap")

set.seed(221)
aucCIFLlogit <- ci.auc(rocFLlogit, method = "bootstrap")
set.seed(221)
aucCIFLrf <- ci.auc(rocFLrf, method = "bootstrap")
set.seed(221)
aucCIAlllogit <- ci.auc(rocAlllogit, method = "bootstrap")
set.seed(221)
aucCIAllrf <- ci.auc(rocAllrf, method = "bootstrap")
```

# Code for Appendix

## The four models with up-sampling within the cross-validation procedure

```{r, results = 'hide'}
tcup <- trainControl(method = "cv", 
                   number = folds,
                   summaryFunction = twoClassSummary, 
                   classProb = TRUE,
                   savePredictions = TRUE,
                   seeds = seeds,
                   index = cvIndex,
                   sampling = "up"
)

# The Fearon/Laitin models.

# Logit
modelFLlogitup <- train(warstds ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil 
                        + nwstate + inst3 + pol4 + ef + relfrac, 
                      metric = "ROC", method = "glm", family = "binomial",
                      trControl = tcup, data = dfTrain
)

summary(modelFLlogitup) 

modelFLlogitup 

confusionMatrix(modelFLlogitup, norm = "average") 


# Random Forest
modelFLrfup <- train(warstds ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil 
                     + nwstate + inst3 + pol4 + ef + relfrac, 
                   metric = "ROC", method = "rf",
                   trControl = tcup, data = dfTrain,
                   keep.inbag = TRUE
)

modelFLrfup

confusionMatrix(modelFLrfup, norm = "average")



############

# Logit
modelAlllogitup <- train(warstds ~ ., 
                       metric = "ROC", method = "glm", family = "binomial",
                       trControl = tcup, data = dfTrain
)

summary(modelAlllogitup) 

modelAlllogitup 

confusionMatrix(modelAlllogitup, norm = "average") 


# Random Forest
modelAllrfup <- train(warstds ~ ., 
                    metric = "ROC", method = "rf",
                    trControl = tcup, data = dfTrain,
                    keep.inbag = TRUE
)

modelAllrfup

confusionMatrix(modelAllrfup, norm = "average")
```


## Confusion Matrices

```{r}
table(pred = predict(modelFLlogitup, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelFLrfup, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAlllogitup, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAllrfup, dfTest, type = "raw"), obs = dfTest$warstds)
```



## Separation Plots and ROC curves for the models with up-sampling

```{r fig.align = 'center', fig.height = 4, fig.width = 7}

separationplot(predict(modelFLlogitup, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="Fearon and Laitin (2003) Logit with up-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)
separationplot(predict(modelFLrfup, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="Fearon and Laitin (2003) Random Forest with up-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)
separationplot(predict(modelAlllogitup, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="All Variables Logit with up-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)
separationplot(predict(modelAllrfup, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="All Variables Random Forest with up-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)

```

```{r}
plot.roc(dfTest$warstds,
         predict(modelFLlogitup, dfTest, type = "prob")$war,
         col = "grey 80", las = 1, xlim = c(1,0), bty = "n", 
         main = "Out-of-sample ROC curves with up-sampling")

plot.roc(dfTest$warstds,
         predict(modelFLrfup, dfTest, type = "prob")$war,
         add = T,
         col = "grey 80", lty = "dashed")


plot.roc(dfTest$warstds,
         predict(modelAlllogitup, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "solid")

plot.roc(dfTest$warstds,
         predict(modelAllrfup, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "dashed")
legend(0.7, 0.25, 
       c(paste("Fearon and Laitin (2003) Logit, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelFLlogitup, dfTest, type = "prob")$war)$auc),2)), 
         paste("Fearon and Laitin (2003) Random Forest, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelFLrfup, dfTest, type = "prob")$war)$auc),2)), 
         paste("All Variables Logit, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelAlllogitup, dfTest, type = "prob")$war)$auc),2)),  
         paste("All Variables Random Forest, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelAllrfup, dfTest, type = "prob")$war)$auc),2))), 
       lty=c("solid", "dashed", "solid", "dashed"), 
       col = c("black","black", "grey 50", "grey 50"), bty="n", 
       cex = .75)
```

## The four models with down-sampling within the cross-validation procedure

```{r The four models using down-sampling, results = 'hide'}
tcdown <- trainControl(method = "cv", 
                   number = folds,
                   summaryFunction = twoClassSummary, 
                   classProb = TRUE,
                   savePredictions = TRUE,
                   seeds = seeds,
                   index = cvIndex,
                   sampling = "down"
)

# The Fearon/Laitin models.

# Logit
modelFLlogitdown <- train(warstds ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil 
                          + nwstate + inst3 + pol4 + ef + relfrac, 
                      metric = "ROC", method = "glm", family = "binomial",
                      trControl = tcdown, data = dfTrain
)

summary(modelFLlogitdown) 

modelFLlogitdown

confusionMatrix(modelFLlogitdown, norm = "average") 


# Random Forest
modelFLrfdown <- train(warstds ~ warhist + ln_gdpen + lpopns + lmtnest + ncontig + oil 
                       + nwstate + inst3 + pol4 + ef + relfrac, 
                   metric = "ROC", method = "rf",
                   trControl = tcdown, data = dfTrain,
                   keep.inbag = TRUE
)

modelFLrfdown

confusionMatrix(modelFLrfdown, norm = "average")



############

# Logit
modelAlllogitdown <- train(warstds ~ ., 
                       metric = "ROC", method = "glm", family = "binomial",
                       trControl = tcdown, data = dfTrain
)

summary(modelAlllogitdown) 

modelAlllogitdown 

confusionMatrix(modelAlllogitdown, norm = "average") 


# Random Forest
modelAllrfdown <- train(warstds ~ ., 
                    metric = "ROC", method = "rf",
                    trControl = tcdown, data = dfTrain,
                    keep.inbag = TRUE
)

modelAllrfdown

confusionMatrix(modelAllrfdown, norm = "average")
```


## Confusion Matrices

```{r}
table(pred = predict(modelFLlogitdown, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelFLrfdown, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAlllogitdown, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAllrfdown, dfTest, type = "raw"), obs = dfTest$warstds)
```



## Separation Plots and ROC curves for the models with down-sampling

```{r fig.align = 'center', fig.height = 4, fig.width = 7}

separationplot(predict(modelFLlogitdown, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="Fearon and Laitin (2003) Logit with down-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)

separationplot(predict(modelFLrfdown, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="Fearon and Laitin (2003) Random Forest with down-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)

separationplot(predict(modelAlllogitdown, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="All Variables Logit with down-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)

separationplot(predict(modelAllrfdown, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2=1,
               show.expected=T, 
               heading="All Variables Random Forest with down-sampling", 
               height=1.5, col0="white", col1="black",newplot = F)

```

```{r}
plot.roc(dfTest$warstds,
         predict(modelFLlogitdown, dfTest, type = "prob")$war,
         col = "grey 80", las = 1, xlim = c(1,0), bty = "n", main = "Out-of-sample ROC curves with down-sampling")

plot.roc(dfTest$warstds,
         predict(modelFLrfdown, dfTest, type = "prob")$war,
         add = T,
         col = "grey 80", lty = "dashed")


plot.roc(dfTest$warstds,
         predict(modelAlllogitdown, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "solid")

plot.roc(dfTest$warstds,
         predict(modelAllrfdown, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "dashed")
legend(0.7, 0.25, 
       c(paste("Fearon and Laitin (2003) Logit, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelFLlogitdown, dfTest, type = "prob")$war)$auc),2)), 
         paste("Fearon and Laitin (2003) Random Forest, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelFLrfdown, dfTest, type = "prob")$war)$auc),2)), 
         paste("All Variables Logit, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelAlllogitdown, dfTest, type = "prob")$war)$auc),2)),  
         paste("All Variables Random Forest, AUC = ", 
               round(as.numeric(roc(dfTest$warstds,
                                    predict(modelAllrfdown, dfTest, type = "prob")$war)$auc),2))), 
       lty=c("solid", "dashed", "solid", "dashed"), 
       col = c("black","black", "grey 50", "grey 50"), bty="n", 
       cex = .75)
```


## Collier/Hoeffler Logit
```{r The Collier/Hoeffler Logit model, results = 'hide'}
modelCHlogit <- train(warstds ~ sxpnew+sxpsq+ln_gdpen+gdpgrowth+warhist+lmtnest+ef
                      +popdense+lpopns+coldwar+seceduc+ptime, 
                      metric = "ROC", method = "glm", family = "binomial",
                      trControl = tc, data = dfTrain
                      )

summary(modelCHlogit) # Not in paper

modelCHlogit # Not in paper

confusionMatrix(modelCHlogit, norm = "average") # Not in paper
```

### Collier/Hoeffler Random Forest
```{r The Random Forest model using the Collier/Hoeffler variables, results = 'hide'}
modelCHrf <- train(warstds ~ sxpnew+sxpsq+ln_gdpen+gdpgrowth+warhist+lmtnest+ef+popdense
                  +lpopns+coldwar+seceduc+ptime, 
                   metric = "ROC", method = "rf",
                   trControl = tc, data = dfTrain,
                   keep.inbag = TRUE
                   )

modelCHrf # Not in paper

confusionMatrix(modelCHrf, norm = "average") # Not in paper
```


## The plots and confusion matrices to compare the models. Collier/Hoeffler vs. all 90 variables

### Confusion Matrices for the models at a threshold for positive prediction of 0.5
```{r Confusion Matrices with Collier/Hoeffler specifications}
table(pred = predict(modelCHlogit, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelCHrf, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAlllogit, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAllrf, dfTest, type = "raw"), obs = dfTest$warstds)
```

### Separation Plots for the four models (Collier/Hoeffler)
```{r Separation Plots for the models with the Collier/Hoeffler variables, fig.align = 'center', fig.height = 4, fig.width = 7}

separationplot(predict(modelCHlogit, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "Collier and Hoeffler Variables (2004) Logit", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelCHrf, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "Collier and Hoeffler Variables (2004) Random Forest", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelAlllogit, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "All Variables Logit", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelAllrf, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "All Variables Random Forest", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)
```

### ROC curves for the four models (Collier/Hoeffler)
```{r ROC curves for the Collier/Hoeffler specifications}
plot.roc(dfTest$warstds,
         predict(modelCHlogit, dfTest, type = "prob")$war,
         col = "grey 80", las = 1, xlim = c(1,0), bty = "n", 
         main = "Out-of-sample ROC curves")

plot.roc(dfTest$warstds,
         predict(modelCHrf, dfTest, type = "prob")$war,
         add = T,
         col = "grey 80", lty = "dashed")


plot.roc(dfTest$warstds,
         predict(modelAlllogit, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "solid")

plot.roc(dfTest$warstds,
         predict(modelAllrf, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "dashed")
legend(0.5, 0.25, c("Collier and Hoeffler (2004) Logit", 
                    "Collier and Hoeffler (2004) Random Forest", 
                     "All Variables Logit",
                    "All Variables Random Forest" ), 
       lty=c("solid", "dashed", "solid", "dashed"), 
       col = c("grey 80","grey 80", "grey 50", "grey 50"), bty="n", 
       cex = .75)
```

## Hegre/Sambanis model specifications

### Logit
```{r The Hegre/Sambanis Logit model, results = 'hide'}
modelHSlogit <- train(warstds ~ lpopns+ln_gdpen+inst3+parreg+geo34+proxregc+gdpgrowth
                      +anoc+partfree+nat_war+lmtnest+decade1+pol4sq+nwstate
                      +regd4_alt+etdo4590+milper+geo1+tnatwar+presi, 
                      metric = "ROC", method = "glm", family = "binomial",
                      trControl = tc, data = dfTrain
                      )

summary(modelHSlogit) # Not in paper

modelHSlogit # Not in paper

confusionMatrix(modelHSlogit, norm = "average") # Not in paper
```

### Random Forest
```{r The Random Forest model using the Hegre/Sambanis variables, results = 'hide'}
modelHSrf <- train(warstds ~ lpopns+ln_gdpen+inst3+parreg+geo34+proxregc+gdpgrowth+anoc
                   +partfree+nat_war+lmtnest+decade1+pol4sq+nwstate+regd4_alt+etdo4590
                   +milper+geo1+tnatwar+presi, 
                   metric = "ROC", method = "rf",
                   trControl = tc, data = dfTrain,
                   keep.inbag = TRUE
                   )

modelHSrf # Not in paper

confusionMatrix(modelHSrf, norm = "average") # Not in paper
```


## The plots and confusion matrices to compare the models. Hegre/Sambanis vs. all 90 variables

### Confusion Matrices for the models at a threshold for positive prediction of 0.5
```{r Confusion Matrices with Hegre/Sambanis specifications}
table(pred = predict(modelHSlogit, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelHSrf, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAlllogit, dfTest, type = "raw"), obs = dfTest$warstds)
table(pred = predict(modelAllrf, dfTest, type = "raw"), obs = dfTest$warstds)
```

### Separation Plots for the four models (Hegre/Sambanis)
```{r Separation Plots for the models with the Hegre/Sambanis variables, fig.align = 'center', fig.height = 4, fig.width = 7}

separationplot(predict(modelHSlogit, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "Hegre and Sambanis Variables (2006) Logit", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelHSrf, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "Hegre and Sambanis Variables (2006) Random Forest", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelAlllogit, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "All Variables Logit", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)

separationplot(predict(modelAllrf, dfTest, type = "prob")$war, 
               as.numeric(dfTest$warstds)-1, type = "line", 
               line = T, lwd2 = 1,
               show.expected = T, 
               heading = "All Variables Random Forest", 
               height = 1.5, col0 = "white", col1 = "black", newplot = F)
```

### ROC curves for the four models (Hegre/Sambanis)
```{r ROC curves for the Hegre/Sambanis specifications}
plot.roc(dfTest$warstds,
         predict(modelHSlogit, dfTest, type = "prob")$war,
         col = "grey 80", las = 1, xlim = c(1,0), bty = "n", 
         main = "Out-of-sample ROC curves")

plot.roc(dfTest$warstds,
         predict(modelHSrf, dfTest, type = "prob")$war,
         add = T,
         col = "grey 80", lty = "dashed")


plot.roc(dfTest$warstds,
         predict(modelAlllogit, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "solid")

plot.roc(dfTest$warstds,
         predict(modelAllrf, dfTest, type = "prob")$war,
         add = T,
         col = "grey 50", lty = "dashed")
legend(0.5, 0.25, c("Hegre and Sambanis (2006) Logit", 
                    "Hegre and Sambanis (2006) Random Forest", 
                    "All Variables Logit", 
                    "All Variables Random Forest" ), 
       lty=c("solid", "dashed", "solid", "dashed"), 
       col = c("grey 80","grey 80", "grey 50", "grey 50"), bty="n", 
       cex = .75)
```


